---
title: "SDB stranding and water temp"
output: html_notebook
---

Looking at relationship between the number of strandings and SDB water temperature. This analysis is for Robin LeRoux - stranding meeting in October 2018

```{r}
rm(list = ls())
.libPaths("C:/Users/tomo.eguchi/AppData/Local/R")
ifelse(Sys.info()[1] == 'Linux',
       source('~/Documents/R/tools/TomosFunctions.R'),
       source('~/R/tools/TomosFunctions.R'))
library(dplyr)
library(ggplot2)

data1 <- read.csv("data/TurtlesWaterTemp_Jan2019a.csv")
data1 %>% mutate(WaterTempC = F2C(WaterTemp),
                 WaterTempC2 = WaterTempC^2,
                 WaterTempC_lag = F2C(WaterTemp_lag),
                 WaterTempC2_lag = WaterTempC_lag^2) -> data1

p.1 <- ggplot(data = data1) + 
  geom_point(aes(x = WaterTempC, y = Stranding),
             size = 1.8,
             color = "red")

p.2 <- ggplot(data = data1) + 
  geom_line(aes(x = MonthNum, y = Stranding))

p.3 <- ggplot(data = data1) + 
  geom_line(aes(x = MonthNum, y = WaterTempC))

fit.1 <- lm(Stranding ~ WaterTempC, data = data1)

```

Looking at the relationship between water temp and stranding, there is a pretty good relationship but maybe not linear? 
```{r}
print(p.1)
```

A linear regression says so too.

```{r}
summary(fit.1)
data1 %>% mutate(linear = predict(fit.1)) -> data1
```

```{r}
p.1.2 <- p.1 + geom_line(data = data1, 
                         aes(x = WaterTempC, y = linear))
print(p.1.2)
```

How about a non-linear regression with a polynomial?

```{r}
#data1 %>% mutate(WaterTempC2 = WaterTempC * WaterTempC) -> data1
fit.1.1 <- lm(Stranding ~ WaterTempC + WaterTempC2, data = data1)
pred.fit.1 <- predict(fit.1.1, 
                      newdata = data.frame(WaterTempC = seq(from = min(data1$WaterTempC),
                                                           to = max(data1$WaterTempC),
                                                           by = 0.1), 
                                           WaterTempC2 = (seq(from = min(data1$WaterTempC),
                                                             to = max(data1$WaterTempC),
                                                             by = 0.1)^2)),
                      se.fit = T,
                      interval = "confidence")

pred.data.1 <- data.frame(WaterTempC = seq(from = min(data1$WaterTempC),
                                           to = max(data1$WaterTempC),
                                           by = 0.1),
                          fit = pred.fit.1$fit[,1],
                          lower = pred.fit.1$fit[,2],
                          upper = pred.fit.1$fit[,3]) 

summary(fit.1.1)
```

A little improvements in the fit. 

```{r}
p.1.3 <- p.1 + 
  geom_line(data = pred.data.1, 
            aes(x = WaterTempC, y = fit),
            color = "red",
            size = 1.2) + 
  geom_line(data = pred.data.1,
            aes(x = WaterTempC, y = lower), 
            color = "red", linetype = 2) + 
  geom_line(data = pred.data.1,
            aes(x = WaterTempC, y = upper), color = "red", linetype = 2) + 
  labs(x = expression("Temperature " ( degree*C)), 
                      y = "# Stranding") + 
  theme(axis.text = element_text(size = 12))
print(p.1.3)
ggsave(filename = "figures/tempVsStranding_noLag.png",
       device = "png", dpi = 600,
       plot = p.1.3)

```


When looking at the number of strandings per month, there is an obvious peak.
```{r}
print(p.2)
```

When looking at the water temp in the same way...

```{r}
print(p.3)
```

There is a lag of about a month... So, why not shift data a bit

```{r}
# data2 <- data.frame(MonthNum = data1$MonthNum,
#                     WaterTempC = data1$WaterTempC,
#                     Stranding2 = c(data1$Stranding[2:12], NA),
#                     WaterTempC2 = data1$WaterTempC2)

p.2.1 <- ggplot(data = data1) + 
  geom_point(aes(x = F2C(WaterTemp_lag), 
                 y = Stranding),
             color = "darkblue",
             size = 2)

p.2.2 <- ggplot(data = data1) + 
  geom_line(aes(x = MonthNum, y = Stranding))

p.2.3 <- ggplot(data = data1) + 
  geom_line(aes(x = MonthNum, y = WaterTempC_lag))


```

```{r}
print(p.2.1)
```

The non-linearity persists... so let's just use a polynomial from the beginning.


```{r}
print(p.2.2)
```

Then the water temp:

```{r}
print(p.2.3)
```

The peaks match better.

Now look at the regression:

```{r}
fit.2 <- lm(Stranding ~ WaterTempC_lag + WaterTempC2_lag, 
            data = data1)

pred.fit.2 <- predict(fit.2,
                      newdata = data.frame(WaterTempC_lag = seq(from = min(data1$WaterTempC_lag),
                                                            to = max(data1$WaterTempC_lag),
                                                            by = 0.1), 
                                           WaterTempC2_lag = (seq(from = min(data1$WaterTempC_lag),
                                                              to = max(data1$WaterTempC_lag),
                                                              by = 0.1))^2),
                      se.fit = T,
                      interval = "confidence")

pred.data <- data.frame(WaterTempC = seq(from = min(data1$WaterTempC_lag),
                                         to = max(data1$WaterTempC_lag),
                                         by = 0.1),
                        fit = pred.fit.2$fit[,1],
                        lower = pred.fit.2$fit[,2],
                        upper = pred.fit.2$fit[,3]) 
summary(fit.2)
```

```{r}
p.2.2 <- p.2.1 + 
  geom_line(data = pred.data, 
            aes(x = WaterTempC, y = fit),
            size = 1.3,
            color = "darkblue") + 
  geom_line(data = pred.data,
            aes(x = WaterTempC, y = lower), 
            linetype = 2,
            size = 1,
            color = "darkblue") + 
  geom_line(data = pred.data,
            aes(x = WaterTempC, y = upper), linetype = 2,
            size = 1,
            color = "darkblue") + 
  labs(x = expression("Temperature " ( degree*C)), 
                      y = "# Stranding") + 
  theme(axis.text = element_text(size = 12))

print(p.2.2)  
ggsave(filename = "figures/WaterTempVsStrandings.png", 
       plot = p.2.2, dpi = 600,
       device = "png")
```

